Co-moving stars in \textsl{Gaia DR1}

### 2016-11-17 paper/ms.tex
Minor typo fix and edits

At small separations ($0.001-1$~pc), they correspond to wide binaries (and
multiples) that are either weakly gravitationally bound or currently being
dispersed.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived their dynamic birth environment which
provides a window to the star formation process.
They may also be disrupted by Galactic tides, or other massive objects
such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models of these two types as the
two stars were likely born at the same time with the same chemical compositions,
but evolve indepdently becayse if their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

@@

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr < 8$,  to \tgas,
which leaves 619,618 stars.
For each surviving star, we then establish an initial sample of possible
co-moving partners by selecting all other stars within 10 pc with a difference
in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for this position cut and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\approx$271k initial pairs examined.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of likelihood ratios for
slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
\todo{Fix in below and in the figure, $\Delta v_t$ should be $|\Delta v_t|$}
Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
their likelihood distribution is narrowly peaked above 6.
The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
and the number of random pairs increasingly dominate.
On the right panel of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on $\Delta v_t$ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

>>>>

At small separations ($0.001-1$~pc), they correspond to wide binaries (and
multiples) that are either weakly gravitationally bound or currently being
dispersed.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived their dynamic birth environment which
provides a window to the star formation process.
They may also be disrupted by Galactic tides, or other massive objects
such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models of these two types as the
two stars were likely born at the same time with the same chemical compositions,
but evolved indepdently because of their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

@@

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to \tgas,
which leaves 619,618 stars.
For each surviving star, we then establish an initial sample of possible
co-moving partners by selecting all other stars within 10 pc with a difference
in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for this position cut and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k initial pairs examined.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of likelihood ratios for
slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
\todo{Fix in below and in the figure, $\Delta v_t$ should be $|\Delta v_t|$}
Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
their likelihood distribution is narrowly peaked above 6.
The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
and the number of random pairs increasingly dominate.
On the right panel of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on $\Delta v_t$ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.
